\Large{All in the family:\\German twin finishing times in\\the 2016 women's Olympic marathon}

### 2017-02-24 paper/chance_memo.tex
misc edits
updated histograms, simulations code, paper text, and memo text

  The $p$-value for the quadratic term is -1.88, which is not
  particularly compelling.  We have accordingly dropped the quadratic
  terms in the scatter plots on p.\ 4 and, implicitly, from the
  Studentized residual plots on p.\ 6.
  %% Note: xr labels are
  %% breaking; looks like some sort of
  %% label clash with another package.
  %% Hardcoding here is poor style.

@@

  We have added age to our simulation model, and the results are
  qualitatively equivalent.  In addition, if one adds age to the
  linear specification depicted in our basic scatter plot (now without
  a quadratic term, as noted above), age is a significant predictor of
  marathon finishing time, even controlling for personal best
  ($p \approx 2.426$).  Older runners are slower, \emph{ceteris
    paribus}.

@@

@@

>>>>

  The $t$-value for the quadratic term is -1.88 and the corresponding $p$-value is .062, which is not
  particularly compelling.  We have accordingly dropped the quadratic
  terms in the scatter plots on p.\ 4 and, implicitly, from the
  Studentized residual plots on p.\ 6.
  %% Note: xr labels are
  %% breaking; looks like some sort of
  %% label clash with another package.
  %% Hardcoding here is poor style.

@@

  We have added age to our simulation model, and the results are
  qualitatively equivalent.  In addition, if one adds age to the
  linear specification depicted in our basic scatter plot (now without
  a quadratic term, as noted above), age is a significant predictor of
  marathon finishing time, even controlling for personal best
  ($t \approx 2.426$, $p \approx 0.017$).  Older runners are slower, \emph{ceteris
    paribus}.

@@

  We have highlighted the left side bins in red and annotated the plot 
  with a description of the size of the those bins. This emphasizes the finishes
  that are the main interest of the result.

@@

  We agree that this is hard to see.  The text now points out that bin
  width is 30 seconds.

### 2017-02-24 paper/paper.tex
updated histograms, simulations code, paper text, and memo text

From our simulations we can generate intervals which describe the
extent of the variability in marathon finishing times. For example, in
95\% of the simulations in which Anna Hahner completed the marathon,
her finishing time was between 9,480 seconds and 10,703 seconds. This
is consistent with Anna's observed finishing time in Rio, which was
9,932 seconds. In fact, our simulations estimate that Anna should have
finished the Rio marathon in 10,092 seconds, which is slightly more
than two and one-half minutes slower than her actual time. Lisa
Hahner's corresponding 95\% interval is 9,427 to 10,619 with an
observed finishing time of 9,933 seconds.

The bottom line here is that our simulated marathon finishes are
consistent with the Hahner twins' finishing times. And note that the
regression model underlying the simulations was estimated without the
Hahner twins (and same for the Kim twins and Luik triplets). According
to our simulations, then, neither Hahner twin ran appreciably slowly,
conditional on personal best times prior to the Rio Olympics.

% 95% simulated interval for Anna Hahner finishing time in seconds: 9480 10703 
% 95% simulated interval for Lisa Hahner finishing time in seconds: 9427 10619

@@

The histograms in Figure \ref{fig:simdiff} raise questions about the
credibility of Anna and Lisa Hahner's story and in particular suggest
that a simultaneous finish in the Rio marathon would be very rare if
Anna and Lisa had run independently of each other. For example, in
fewer than 300 of the 10,000 simulated races did Anna and Lisa Hahner
finish within 30 seconds of one another. Moreover, they finished in
consecutive rank in fewer than 175 of the 10,000 simulated races.  The
close finish that was observed, where Anna and Lisa crossed the finish
line one after the other, would have been highly unlikely if the two
German twins had raced independently of each other.

>>>>

From our simulations we can generate intervals which describe the
extent of the variability in marathon finishing times. For example, in
95\% of the simulations in which Anna Hahner completed the marathon,
her finishing time was between 8,369 seconds and 9,999 seconds. This
is consistent with Anna's observed finishing time in Rio, which was
9,932 seconds. In fact, our simulations estimate that Anna should have
finished the Rio marathon in 9,181 seconds, which is slightly more
than 12 minutes faster than her actual time. Lisa
Hahner's corresponding 95\% interval is 8,507 to 10,143 with an
observed finishing time of 9,933 seconds.

The bottom line here is that our simulated marathon finishes are
consistent with the Hahner twins' finishing times. And note that the
regression model underlying the simulations was estimated without the
Hahner twins (and same for the Kim twins and Luik triplets). According
to our simulation, then, both Hahner twins ran appreciably slowly,
conditional on personal best times prior to the Rio Olympics and age.

% 95% simulated interval for Anna Hahner finishing time in seconds: 8369 9999 
% 95% simulated interval for Lisa Hahner finishing time in seconds: 8507 10143

% > mean(hahner_df_simwithage$anna_final, na.rm = T)
% [1] 9181.137
% > mean(hahner_df_simwithage$lisa_final, na.rm = T)
% [1] 9319.427

% > (9932 - 9181.137)/60
% [1] 12.51438

@@

The histograms in Figure \ref{fig:simdiff} raise questions about the
credibility of Anna and Lisa Hahner's story and in particular suggest
that a simultaneous finish in the Rio marathon would be very rare if
Anna and Lisa had run independently of each other. For example, in
fewer than 300 of the 10,000 simulated races did Anna and Lisa Hahner
finish within 30 seconds of one another. Moreover, they finished in
consecutive rank in fewer than 200 of the 10,000 simulated races.  The
close finish that was observed, where Anna and Lisa crossed the finish
line one after the other, would have been highly unlikely if the two
German twins had raced independently of each other.
